---
title: "SWBM usage"
author: "Beni Stocker"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(rsofun)
library(dplyr)
library(ggplot2)
library(tidyr)
library(recipes)
# library(FluxDataKit)
library(readr)
library(stringr)

source("R/fill_missing_netrad.R")

# fake variable as optimization isn't used
pars <- list()
pars$par["kphio"] <- 0.05
```

## Net radiation as forcing

### Read data

Check whether sufficient sites have net radiation data.
```{r}
filnam <- "F:/master thesis/data/df_fluxnet.rds"
if (!file.exists(filnam)){
  df_fluxnet <- ingest(
    siteinfo  = ingestr::siteinfo_fluxnet2015,
    source    = "fluxnet",
    getvars   = list(
      netrad = "NETRAD",
      ppfd = "SW_IN_F",
      temp = "TA_F_DAY"
      ),
    dir       = "~/data/FLUXNET-2015_Tier1/20191024/DD/",
    timescale = "d",
    settings  = list(
      getswc = TRUE,
      threshold_NETRAD = 0.8
      ),
    )
  saveRDS(df_fluxnet, file = filnam)
} else {
  df_fluxnet <- readRDS(filnam)
}
```

Missing data across all sites.
```{r}
df_fluxnet |> 
  tidyr::unnest(data) |> 
  select(netrad) |> 
  visdat::vis_miss(warn_large_data = FALSE, cluster = FALSE)
```

Get fraction of missing data per site.
```{r}
calc_f_missing <- function(df, varnam){
  vec <- df |> 
    pull(!!varnam)
  (sum(is.na(vec)) / nrow(df))
}

df_missing <- df_fluxnet |> 
  mutate(f_missing = purrr::map_dbl(data, ~calc_f_missing(., "netrad")))
```

Distribution of missing data.
```{r}
df_missing |> 
  ggplot(aes(x = f_missing, y = ..count..)) +
  geom_histogram()
```

### Site selection

Available sites with maximum 10% missing.
```{r}
df_sites <- df_missing |> 
  select(-data) |> 
  filter(f_missing <= 0.1) |> 
  arrange(desc(f_missing)) |> 
  left_join(
    ingestr::siteinfo_fluxnet2015,
    by = "sitename"
  ) |> 
  filter(classid != "CRO") |> 
  mutate(nyears = as.numeric(year_end) - as.numeric(year_start) + 1)
```

Selecting 17 (to have enough to start with):

```{r}
df_sites <- df_sites |> 
  filter(sitename %in% c("FR-LBr", "FR-Pue", "US-Ton", "US-NR1", "NL-Loo", "US-MMS", "US-Wkg", "AT-Neu", "AU-DaS", "DE-Tha", "GF-Guy", "US-NR1", "US-UMB", "AU-Tum", "US-Me2", "DE-Hai", "US-SRM"))
```

### Fill missing data

Fill missing data using KNN.
```{r}
df_fluxnet_filled <- df_fluxnet |> 
  filter(sitename %in% df_sites$sitename) |> 
  mutate(data = purrr::map(data, ~fill_missing_netrad(.)))
```

With filled values.
```{r}
df_fluxnet_filled |> 
  filter(sitename == "FR-Pue") |> 
  unnest() |>
  ggplot() +
  geom_point(aes(date, netrad_filled), color = "red") +
  geom_point(aes(date, netrad))
```

Table of missing data
```{r}
df_missing |> 
  select(sitename, f_missing) |> 
  knitr::kable()
```


## SWBM in rsofun

The Simple Water Bucket Model by Orth et al. (2013) is implemented in rsofun and uses net radiation as a forcing (as opposed to SPLASH, where it is calculated as a function of (prescribed) solar radiation, temperature, and cloud cover). 

To use it, check out rsofun branch `rwsc` and compile it.

### Simulation parameters

By default, net radiation is not read as an input. Change this to net radiation treated as a model forcing.

```{r}
p_model_drivers$params_siml[[1]]$in_netrad <- TRUE
```

### Forcing data

Get the FLUXNET 2015 forcing data from sofunCalVal.
```{r}
load("F:/sofunCalVal_jr/data/df_drivers_fluxnet2015_allsites.rda")  # loads 'df_drivers_fluxnet2015_allsites'
```

Subset sites.
```{r}
df_drivers <- df_drivers_fluxnet2015_allsites |> 
  filter(sitename %in% df_sites$sitename)
```

Complement forcing with filled net radiation data.

```{r}
df_drivers <- df_drivers |> 
  select(sitename, forcing) |> 
  unnest(forcing) |> 
  left_join(
    df_fluxnet_filled |> 
      unnest(data) |> 
      select(date, netrad = netrad_filled),
    by = c("sitename", "date")
  ) |> 
  group_by(sitename) |> 
  nest() |> 
  rename(forcing = data) |> 
  right_join(
    df_drivers |> 
      select(-forcing),
    by = "sitename"
  )
```

### Validation data

Adapted from https://github.com/computationales/sofunCalVal/blob/main/analysis/00_prepare_fluxnet_driver_data.R:

Get data from FluxDataKit (downloaded from Zenodo to `~/data/flux_data_kit_beta/fluxes/`)
```{r}
filnam <- "F:/master thesis/data/df_fdk.rds"

if (!file.exists(filnam)){
  # # create CSV files with daily data
  # list_path <- purrr::map(
  #   as.list(df_sites$sitename),
  #   ~try(fdk_convert_lsm(
  #     site = .,
  #     path = "~/data/flux_data_kit_beta/fluxes",
  #     fluxnet_format = TRUE,
  #     out_path = "~/data/flux_data_kit_beta/fluxes_csv"
  #     )  
  #   ))
  # 
  # # subset to only those for which FDK data is available (12)
  # paths <- list_path[purrr::map_chr(list_path, ~class(.)) == "character"]
  
  paths <- list.files("~/data/flux_data_kit_beta/fluxes_csv", 
                      pattern = "DD",
                      full.names = TRUE)
  
  # read data and convert to data frame
  list_fdk <- purrr::map(
    paths,
    ~read_csv(.)
  )
  names(list_fdk) <- paths
  df_fdk <- bind_rows(
    list_fdk,
    .id = "path"
    ) |> 
    mutate(sitename = str_sub(path, start = 63, end = 68))
  
  write_rds(df_fdk, file = filnam)
  
} else {
  
  df_fdk <- readRDS(filnam)
  
}
```

Get LE data and retain only good quality data (50% cutoff leaves enough data)
```{r}
df_obs <- df_fdk |> 
  filter(sitename %in% df_sites$sitename) |> 
  select(sitename, date = TIMESTAMP, LE_F_MDS, LE_F_MDS_QC) |> 
  mutate(le = ifelse(LE_F_MDS_QC > 0.5, LE_F_MDS, NA))
```

Lots of missing data?
```{r}
df_missing2 <- df_obs |> 
  group_by(sitename) |> 
  nest() |> 
  mutate(f_missing = purrr::map_dbl(data, ~calc_f_missing(., "le"))) |> 
  select(-data)

df_missing2 |> 
  ggplot(aes(x = f_missing, y = ..density..)) +
  geom_histogram()
```

### Uncalibrated run

This is now done with an rsofun version where net radiation can be used as a forcing. This is available in branch `rwsc`. Install rsofun from that branch and run it below using default model parameters.

```{r}
# simulation parameter needs to be re-defined.
df_drivers$params_siml[[1]]$in_netrad <- TRUE

df_drivers2 <- df_drivers |> 
  
  # retain only those for which we have observational data from fdk
  filter(sitename %in% unique(df_obs$sitename)) |> 
  
  rename(params_soil = df_soiltexture,
         site_info = siteinfo) |> 
  
  # add missing variables to forcing (not actually used here)
  mutate(forcing = purrr::map(forcing, 
                              ~mutate(., 
                                      tmin = temp, 
                                      tmax = temp)))

# use default parameters
params_modl <- list(
    kphio           = 0.09423773,
    soilm_par_a     = 0.33349283,
    soilm_par_b     = 1.45602286,
    tau_acclim_tempstress = 10,
    par_shape_tempstress  = 0.0,
    exp_et = 0.06,
    beta_et = 0.66,
    rwsc = 200
  )

# run the model for these parameters
# for a single site
# output <- runread_pmodel_f(
#   df_drivers2 |>
#     filter(sitename == "FR-Pue"),
#   par = params_modl
#   )

output <- runread_pmodel_f(
  df_drivers2|>
  filter(sitename == "FR-Pue"),
  par = params_modl
  )
```

### Plot

Plot modelled and observed net radiation
```{r}
df_drivers2 |> 
  filter(sitename == "FR-Pue") |> 
  unnest(forcing) |> 
  select(date, netrad_obs = netrad) |> 
  left_join(
    output$data[[1]],
    by = "date"
  ) |> 
  ggplot() +
  geom_line(aes(date, netrad_obs), color = "red") 
# +
#   geom_line(aes(date, netrad/(24*60*60)), color = "red")
```

Plot modelled PET and AET
```{r}
output$data[[1]] |>
  ggplot() + 
  geom_line(aes(date, pet), color = "royalblue") +
  geom_line(aes(date, transp))
```

Plot observed LE.
```{r}
tmp <- df_obs |> 
  filter(sitename == "FR-Pue") |> 
  left_join(
    output$data[[1]],
    by = "date"
  ) |> 
  mutate(latenth = latenth / (24 * 60 * 60) ) |> 
  left_join(
    df_fluxnet |> 
      filter(sitename == "FR-Pue") |> 
      unnest(data) |> 
      select(date, starts_with("SWC_")),
    by = c("sitename", "date")
  )

# out <- tmp |> 
#   rbeni::analyse_modobs2("latenth", "LE_F_MDS")
# out$gg

tmp |> 
  ggplot() +
  geom_line(aes(date, latenth), color = "red") +
  geom_point(aes(date, LE_F_MDS))
```

Plot soil moisture.
```{r}
tmp |> 
  ggplot() +
  geom_line(aes(date, wscal), color = "royalblue") +
  geom_line(aes(date, SWC_F_MDS_1), color = "grey70") +
  geom_hline(yintercept = 1, linetype = "dotted") +
  ylim(0, 1.2) 
```

### Calibrate

Calibratable parameters:

- `exp_et` $\gamma$ in Orth et al., 2013
- `beta` $\beta_0$ in Orth et al., 2013
- `rzwsc` $c_s$ in Orth et al., 2013

Specify calibration settings.
```{r}
source("R/cost_mae_swbm.R")

settings <- list(
  method  = "gensa",
  metric  = cost_mae_swbm,
  control = list(max.call = 1e3),  # change this to 1e4 for final calibration
  par     = list(
    exp_et  = list(lower=0.01, upper=0.5, init = 0.06),
    beta_et = list(lower=0.1,  upper=1.0, init = 0.66),
    rwsc    = list(lower=10,   upper=1000, init = 200)
    )
)
```

Run calibration.
```{r}
pars <- calib_sofun(
    drivers = df_drivers2 |> 
      filter(sitename == "FR-Pue"),  
    obs = df_obs |> 
      mutate(le = le*24*60*60) |> 
      group_by(sitename) |> 
      nest() |> 
      filter(sitename == "FR-Pue"),
    settings = settings
  )
```

Run model with optimized parameters.
```{r}
params_modl <- list(
    kphio           = 0.09423773,
    soilm_par_a     = 0.33349283,
    soilm_par_b     = 1.45602286,
    tau_acclim_tempstress = 10,
    par_shape_tempstress  = 0.0,
    exp_et = pars$par[["exp_et"]],
    beta_et = pars$par[["beta_et"]],
    rwsc = pars$par[["rwsc"]]
  )

# run the model for these parameters
output <- runread_pmodel_f(
  df_drivers2 |> 
    filter(sitename == "FR-Pue"),
  par = params_modl
  )
```
```{r}
tmp <- df_obs |> 
  filter(sitename == "FR-Pue") |> 
  left_join(
    output$data[[1]],
    by = "date"
  ) |> 
  mutate(latenth = latenth / (24 * 60 * 60) ) |> 
  left_join(
    df_fluxnet |> 
      filter(sitename == "FR-Pue") |> 
      unnest(data) |> 
      select(date, starts_with("SWC_")),
    by = c("sitename", "date")
  )


# visulize it 
tmp |> 
  ggplot() +
  geom_line(aes(date, latenth), color = "red") +
  geom_point(aes(date, LE_F_MDS))
```

Visualize modeled vs observed.
```{r}
tmp <- df_obs |> 
  filter(sitename == "AT-Neu") |> 
  left_join(
    output$data[[1]],
    by = "date"
  ) |> 
  mutate(latenth = latenth / (24 * 60 * 60) ) |> 
  left_join(
    df_fluxnet |> 
      filter(sitename == "AT-Neu") |> 
      unnest(data) |> 
      select(date, starts_with("SWC_")),
    by = c("sitename", "date")
  )

# out <- tmp |> 
#   rbeni::analyse_modobs2("latenth", "LE_F_MDS")
# out$gg
```